---
title: "Cases of Data Science PG"
author: "Secary M"
date: "Trimester `r if(as.integer(format(Sys.time(),'%m'))<5){1}else if(as.integer(format(Sys.time(),'%m'))<9){2}else{3}`, `r format(Sys.time(),'%Y')`"
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
    number_sections: true
    toc_depth: 3
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  fig.width = 6, 
  fig.asp = 0.618, 
  out.width = "70%",
  fig.align = "center", 
  root.dir = '../'
)
pacman::p_load(
  bookdown,
  tidyverse, tidymodels, janitor, paletteer, vip, 
  visdat, naniar, corrplot, plotly, ggridges, viridis, magrittr,
  here, visNetwork, glmnet,
  tidytuesdayR, schrute, harrypotter, ISLR, titanic
)
theme_set(theme_bw())
wdpath <- getwd()
```

# DS_CS01_Linear_Regression_Fuel_Economy

## Case study

#### Set up libraries

```{r}
pacman::p_load(
  tidyverse, 
  tidymodels,
  AppliedPredictiveModeling, 
  skimr, 
  janitor, 
  corrplot, 
  vip
)
```

#### Read in data

```{r}
data(FuelEconomy, package = "AppliedPredictiveModeling")
```


#### Pre-processing

###### Convert to tibble

```{r}
cars2010 <- as_tibble(cars2010)
```

###### Clean names and select predictors

```{r}
cars2010 <- clean_names(cars2010)
cars2010 <- 
  cars2010 %>% 
  select(
    eng_displ, 
    num_cyl, 
    FE = fe, 
    num_gears, 
    drive_desc
  )
cars2010
```

###### Recode the drive predictor

```{r}
cars2010 %>% 
  count(drive_desc)
cars2010<- 
  cars2010 %>% 
  mutate(
    drive = case_when(
      str_detect(drive_desc, "Front") ~ "front", 
      str_detect(drive_desc, "Rear") ~ "rear", 
      TRUE ~ "4WD"
    ))
cars2010 %>% 
  count(drive_desc, drive)
cars2010 <- 
  cars2010 %>% 
  select(-drive_desc)
cars2010
```

###### Convert character into factor

```{r}
cars2010 <- 
  cars2010 %>% 
  mutate_if(
    is.character, factor
  )
cars2010
```


###### Split into training and testing

```{r}
set.seed(2020)
car_split <- initial_split(cars2010, strata = FE)
car_train <- training(car_split)
car_test <- testing(car_split)
```


#### EDA

###### Overview

```{r}
skim(car_train)
```


###### Look at Fuel efficiency

```{r}
car_train %>% 
  ggplot(aes(FE)) + 
  geom_histogram(col = "black", fill = "orange")
```

###### Bivariate relationships

```{r}
car_train %>% 
  ggplot(aes(eng_displ, FE)) + 
  geom_point()
```

```{r}
car_train %>% 
  ggplot(aes(factor(num_cyl), FE)) + 
  geom_boxplot()
```

```{r}
car_train %>% 
  ggplot(aes(factor(num_gears), FE)) + 
  geom_boxplot()
```

```{r}
car_train %>% 
  ggplot(aes(drive, FE)) + geom_boxplot()
```

###### Relationship between predictors

```{r}
car_train %>% 
  select_if(is.numeric) %>% 
  cor() %>% 
  corrplot()
```


###### Preprocessing with recipes

```{r}
car_recipe <- 
  recipe(FE ~ ., data = car_train) %>% 
  step_poly(eng_displ, degree = 2, options = list(raw = TRUE))
car_recipe
```


```{r}
car_recipe <- car_recipe %>% prep()
car_recipe
```


```{r}
car_train <- car_recipe %>% juice()
car_train
```


```{r}
car_test <- car_recipe %>% bake(car_test)
car_test
```

#### Fit model

```{r}
car_lm <- 
  linear_reg() %>% 
  set_engine("lm") %>% 
  fit(FE ~ ., data = car_train)
car_lm
```


#### Use model

###### Model fit

```{r}
car_train %>% 
  add_column(
    predict(car_lm, new_data = car_train)
  ) %>% 
  yardstick::rmse(truth = FE, estimate = .pred)
```

```{r}
car_test %>% 
  add_column(
    predict(car_lm, new_data = car_test)
  ) %>% 
  yardstick::rmse(truth = FE, estimate = .pred)

```

###### Check assumptions

```{r}
car_res <- augment(car_lm$fit)
car_res
```

```{r}
car_res %>% 
  ggplot(aes(.fitted, .resid)) + 
  geom_point() + 
  geom_hline(yintercept = 0)
```

```{r}
car_res %>% 
  ggplot(aes(sample = .resid)) + 
  stat_qq() + 
  stat_qq_line()
```

```{r}
car_res %>% 
  ggplot(aes(.hat, .std.resid)) +
  geom_vline(size = 2, colour = "white", xintercept = 0) +
  geom_hline(size = 2, colour = "white", yintercept = 0) +
  geom_point(aes(size = .cooksd)) + geom_smooth(se = FALSE)
```

###### Variable importance

```{r}
vi(car_lm)
```

```{r}
vip(car_lm, 
    include_type = TRUE, 
    geom = 'point',
    mapping = aes_string(col = "Sign"), size = 5) + 
  scale_color_brewer(palette = "Set1")
```

###### Inference

```{r}
tidy(car_lm$fit)
```

```{r}
tidy(car_lm$fit) %>% 
  arrange(p.value)
```

###### Prediction


```{r}
new_data <- tibble(
  num_cyl = 4, 
  num_gears = 6, 
  eng_displ = 5, 
  drive = "front"
) 
new_data
```

```{r}
new_data <- bake(car_recipe, new_data = new_data)
new_data
```

```{r}
predict(car_lm, new_data = new_data)
predict(car_lm, new_data = new_data, type = "pred_int")
```


# DS_CS02_Data_Cleaning_Eurovision

## Rule 1: Look at the Data

## Rule 2: Is There a Package to Deal With That

```{r}
songs <- jsonlite::fromJSON(glue::glue("{wdpath}/data/eurovision-lyrics.json"))
```

## Rule 3: Break it Into Pieces

```{r}
songs[[1]]
```

```{r}
parse_song <- function(song){
  country <- song$Country
  artist <- song$Artist
  title <- song$Song
  year <- song$Year
  lyrics <- song$Lyrics
  translation <- song$`Lyrics translation`
  place <- song$Pl.
  tibble(
    country, artist, title, year, lyrics, translation, place
  )
}
parse_song(songs[[1]])
```

## Rule 4: Get it Into a Tibble

```{r}
songs <- 
  songs %>% 
  map_df(parse_song, .id = "ID")
songs
```

## Rule 5: Look at Each Column

We need to make sure all of our data makes sense in context. You can find the source data at this link: <https://www.kaggle.com/minitree/eurovision-song-lyrics>

#### ID

This is an ID number assigned to each song. We should check that the song IDs are all unique.

```{r}
songs %>% count(ID) %>% filter(n > 1)
```

They are unique, but they're also not useful, so delete them.

```{r}
songs <- 
  songs %>% 
  select(-ID)
songs
```

#### Artist

Are there any artists who appear more than once?

```{r}
songs %>% 
  count(artist) %>%
  arrange(-n)
```

```{r}
songs %>%
  filter(str_detect(artist, "Jed"))
```

```{r}
songs %>%
  filter(country == "Australia")
```

#### Title

```{r}
songs %>%
  count(title) %>% 
  arrange(-n)
```

#### Year

```{r}
range(songs$year)
```

```{r}
songs %>% 
  count(year) %>% 
  ggplot(aes(year, n)) + 
  geom_point() + 
  theme(axis.text.x = element_text(angle = -90, hjust=0))
```

#### Lyrics

```{r}
songs %>% 
  sample_n(5) %>% 
  pull(lyrics) %>% 
  cat(sep = "\n\n\n")
```

#### Translation

```{r}
songs %>% 
  sample_n(5) %>% 
  pull(translation) %>% 
  cat(sep = "\n\n\n")
```

#### Place

```{r}
songs %>% count(place)
```

```{r}
length(unique(songs$year))
```

```{r}
songs %>% 
  count(year, place) %>% 
  filter(place == "1") %>% 
  arrange(-n)
```

According to Wikpedia, there were four winners in 1969:

<https://en.wikipedia.org/wiki/Eurovision_Song_Contest>

> On only one occasion have multiple winners been declared in a single contest: in 1969, four countries finished the contest with an equal number of votes and due to the lack of a tie-break rule at the time, all four countries were declared winners.[7][157] 

## Rule 6: You May Need to Go Back

Our original parse function just grabbed `translation` as is. After looking at `translation`, we can see that this variable says `English` if the original song lyrics are in English. English.

## Rule 7: Create New Columns?

#### Decade

```{r}
songs
songs <- 
  songs %>% 
  mutate(
    year = parse_number(year), 
    decade = year - year %% 10
  )
songs %>% 
  ggplot(aes(year, decade)) + 
  geom_point()
```

#### Lyrics

```{r}
songs %>% 
  count(translation) %>%
  arrange(-n)
```

```{r}
songs <- 
  songs %>% 
  mutate(
    english_lyrics = ifelse(
      translation == "English",
      lyrics, translation
    ) 
  )
```

## Rule 7: Save the Data and Write it Up

It's critically important to make detailed notes about any changes you have made when cleaning the data.

```{r}
filenamepath <- glue::glue("{wdpath}/data/{lubridate::today()}-eurovision.rds")
filenamepath
write_rds(songs,filenamepath)
```


# DS_CS03_Pre_processing_Credit_Data

## Get data and basic cleaning
#### Read in data
```{r}
data(credit_data, package = "modeldata")
```

#### Convert to tibble
```{r}
credit_data <- as_tibble(credit_data)
credit_data
```

#### Clean names
```{r}
credit_data <- credit_data %>%
  clean_names()
credit_data
```

## Quick gander
```{r}
skimr::skim(credit_data)
```

## Pre-processing
Consists of three steps: 
- recipe
- prep
- bake / juice

#### Recipe
###### Set up types of variables

```{r}
credit_recipe <- recipe(status ~ ., data = credit_data)
credit_recipe
summary(credit_recipe)
```

###### Add steps
######## Center and scale
```{r}
credit_recipe <- 
  credit_recipe %>% 
  step_center(all_numeric()) %>% 
  step_scale(all_numeric())
credit_recipe
```

######## Skewness
```{r}
get_skewness <- function(x){
  ## Remove missing
  x <- x[!is.na(x)]
  ## Get mean
  x_bar <- mean(x)
  ## Get number of observations
  n <- length(x)
  ## Calculate v
  v <- sum( (x - x_bar)^2 ) / (n - 1)
  ## Calculate skewness
  skewness <- sum( (x - x_bar)^3 ) / ((n - 1) * v^(3/2))
  return(skewness)
}
get_skewness(rnorm(100))
get_skewness(rexp(100))
```

```{r}
credit_data %>%
  summarise(
    across(where(is.numeric), get_skewness)
  )
```

```{r}
credit_recipe <- credit_recipe %>%
  step_BoxCox(all_numeric())
credit_recipe
```

######## Missing data
```{r}
vis_miss(credit_data)
```

```{r}
gg_miss_upset(credit_data)
```

```{r}
credit_recipe <- credit_recipe %>%
  step_impute_mode(all_nominal()) %>%
  step_impute_mean(all_numeric())
credit_recipe
```

######## Near zero
```{r}
credit_recipe <- credit_recipe %>%
  step_nzv(all_predictors())
credit_recipe
```

######## Collinearity
```{r}
credit_data %>% 
  select_if(is.numeric) %>% 
  cor(use = "pairwise.complete.obs") %>% 
  corrplot()
```

```{r}
credit_recipe <- credit_recipe %>%
  step_corr(all_numeric())
credit_recipe
```

######## Dummy
```{r}
credit_recipe <- credit_recipe %>%
  step_dummy(all_nominal(), -status)
credit_recipe
```

```{r}
tidy(credit_recipe)
tidy(credit_recipe, n = 1)
```

#### Prep recipe
```{r}
credit_recipe <- credit_recipe %>%
  prep(data = credit_data)
credit_recipe
```

```{r}
tidy(credit_recipe)
tidy(credit_recipe, n = 1)
```

## Get data
#### Juice
```{r}
credit_recipe %>% juice()
```

#### Bake
```{r}
credit_recipe %>% bake(credit_data)
```


# DS_CS04_Cross_Validation_Titanic

## Load data

```{r}
data("titanic_train", package = "titanic")
titanic <- as_tibble(titanic_train)
titanic
```

## Clean names

```{r}
titanic <- titanic %>% clean_names()
titanic
```

## Convert survived to factor

```{r}
titanic <- titanic %>% 
  mutate_all(factor)
titanic
```

## Split into training and test

```{r}
titanic_split <- initial_split(titanic, strata = survived)
titanic_train <- training(titanic_split)
titanic_test <- testing(titanic_split)
```

```{r}
titanic_train
```
```{r}
titanic_test
```


## Get CV of training

```{r}
titanic_cv <- vfold_cv(titanic_train, v = 5, strata = survived)
titanic_cv
```


## Set up workflow

#### Recipe

```{r}
titanic_recipe <- 
  recipe(survived ~ sex + pclass, data = titanic_train) %>% 
  step_dummy(all_predictors()) %>% 
  step_interact(terms = ~starts_with("sex"):starts_with("pclass"))
```

```{r}
titanic_recipe %>% prep() %>% juice()
```

#### Model

```{r}
titanic_model <- logistic_reg() %>% set_engine("glm")
```

```{r}
titanic_wf <- workflow() %>% 
  add_recipe(titanic_recipe) %>% 
  add_model(titanic_model)
titanic_wf
```

```{r}
titanic_fit <- fit(titanic_wf, titanic_train)
titanic_fit %>% tidy()
```

## Fit CV

```{r}
titanic_fits <- fit_resamples(
  titanic_wf, 
  resamples = titanic_cv, 
  control = control_resamples(save_pred = TRUE)
)
titanic_fits
```


## Get metrics

```{r}
titanic_fits %>% collect_metrics()
```
## ROC curves

```{r}
CV_ROC <- 
  titanic_fits %>% 
  collect_predictions() %>% 
  group_by(id) %>% 
  roc_curve(truth = survived, .pred_1, event_level = "second")
CV_ROC %>% autoplot()
```


## Compare to test

```{r}
titanic_test %>% 
  add_column(
    predict(titanic_fit, new_data = titanic_test, type = "prob")
  ) %>% 
  roc_auc(truth = survived, .pred_1, event_level = "second")
```

```{r}
test_ROC <- titanic_test %>% 
  add_column(
    predict(titanic_fit, new_data = titanic_test, type = "prob")
  ) %>% 
  roc_curve(truth = survived, .pred_1, event_level = "second") %>% 
  add_column(id = "test")
bind_rows(CV_ROC, test_ROC) %>% 
  ggplot(aes(1 - specificity, sensitivity, col = id)) + geom_path() + 
  scale_color_brewer(palette = "Set2")
```




# DS_CS05_Ridge_Regression_The_Office
## Read in data

```{r}
data("theoffice", package = "schrute")
office <- theoffice
office
```

So each row is a line from the script, we also have info about season, episode etc. 

Want to predict rating from how much each character speaks. 

## Characters

How many characters?

```{r}
length(unique(office$character))
```

Too many, so we will collapse down

```{r}
office <- 
  office %>% 
  mutate(
    character = fct_lump_n(character, 20)
  )
unique(office$character)
```

## Cleaning

First count the number of lines for each character. 

```{r}
office_long <- 
  office %>% 
  group_by(season, episode, character) %>% 
  summarise(
    n_lines = n(), 
    rating = mean(imdb_rating)
  )
office_long
```

Now convert to wide form

```{r}
office_wide <- 
  office_long %>% 
  pivot_wider(
    names_from = character, 
    values_from = n_lines, 
    values_fill = 0
  )
office_wide
```

## Split

```{r}
office_split <- initial_split(office_wide, strata = rating)
office_train <- training(office_split)
office_test <- testing(office_split)
```

## Preprocessing

```{r}
office_recipe <- 
  recipe(rating ~ ., data = office_train) 
office_recipe
```

Set ID columns

```{r}
office_recipe <- 
  office_recipe %>% 
  update_role(season,episode, new_role = "ID")
office_recipe
```

Scale and center

```{r}
office_recipe <- 
  office_recipe %>% 
  step_center(all_predictors()) %>% 
  step_scale(all_predictors())
```

Check 

```{r}
office_recipe %>% prep() %>% juice()
office_recipe %>% prep() %>% juice() %>% select(rating)
```

## Fit ridge regression

#### Set model

```{r}
office_model <- 
  linear_reg(mixture = 0, penalty = 0.1) %>% 
  set_engine("glmnet")
```

Note still uses lots of lambdas - see fit. Penalty only used for prediction. 

#### Create workflow

```{r}
office_wf <- 
  workflow() %>% 
  add_recipe(office_recipe) %>% 
  add_model(office_model)
office_wf
```

#### Fit workflow

```{r}
office_ridge <- 
  office_wf %>% 
  fit(office_train)
office_ridge
```

```{r}
p <- office_ridge %>% 
  extract_fit_parsnip() %>% 
  .$fit  %>% 
  tidy() %>% 
  filter(
    term != "(Intercept)", 
    lambda < 100
  ) %>% 
  ggplot(aes(lambda, estimate, col = term)) + 
  geom_line() 
p
```

```{r}
ggplotly(p)
```

## Predict

#### Train RMSE

```{r}
office_train %>% 
  add_column(
    predict(office_ridge, new_data = office_train)
  ) %>% 
  ungroup() %>% 
  rmse(rating, .pred)
```

#### Test RMSE

```{r}
office_test %>% 
  add_column(
    predict(office_ridge, new_data = office_test)
  ) %>% 
  ungroup() %>% 
  rmse(rating, .pred)
```

## VIP

```{r}
office_ridge %>% 
  pull_workflow_fit() %>% 
  vi() %>% 
  ggplot(aes(Importance, fct_reorder(Variable, Importance), fill = Sign)) + 
  geom_col() + 
  labs(y = "Character") + 
  scale_fill_paletteer_d("dutchmasters::milkmaid")
```

```{r}
sessionInfo()
```



# DS_CS06_Lasso_Hitters
## Get the Data

```{r}
data("Hitters", package = "ISLR")
hitters <- as_tibble(Hitters)
hitters
```

## Cleaning

#### Clean Variable Names

```{r}
hitters %<>% janitor::clean_names()
hitters
```

#### Remove NA Values

```{r}
hitters %<>% drop_na()
hitters
```

#### Create Train and Test Splits

```{r}
hitters_split <- initial_split(hitters, strata = salary)
hitters_train <- training(hitters_split)
hitters_test <- testing(hitters_split)
```

##Workflow

#### Recipe

```{r}
hitters_recipe <- recipe(salary ~ ., data = hitters) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_nominal_predictors())
hitters_recipe %>% prep() %>% bake(new_data = NULL)
hitters_recipe %>% prep() %>% bake(new_data = NULL) %>% pull(salary)
```

#### Model

```{r}
hitters_model <- linear_reg(mixture = 1, penalty = tune()) %>%
  set_engine("glmnet") %>%
  set_mode("regression")
hitters_model
```

#### Workflow

```{r}
hitters_wf <- workflow() %>%
  add_model(hitters_model) %>%
  add_recipe(hitters_recipe)
hitters_wf
```

## Tuning

#### Penalty Grid

```{r}
##hitters_grid <- grid_regular(penalty(), levels = 50)
hitters_grid <- tibble(penalty = seq(.5,50,.5))
hitters_grid
```

#### Cross Validation Folds

```{r}
set.seed(2022)
hitters_cv <- vfold_cv(hitters_train, strata = salary)
hitters_cv
```

#### Tune Model (Penalty Term)

```{r}
doParallel::registerDoParallel()
hitters_tune <- tune_grid(object = hitters_wf,
                          resamples = hitters_cv,
                          grid = hitters_grid)
hitters_tune %>% collect_metrics()
```

## Decide on a Model

```{r}
p <- hitters_tune %>%
  collect_metrics() %>%
  filter(.metric == "rmse") %>%
  ggplot(aes(penalty, mean)) +
  geom_point() +
  geom_errorbar(aes(ymin = mean - std_err, ymax = mean + std_err))
p
```

```{r}
plotly::ggplotly(p)
```

```{r}
show_best(hitters_tune, metric = "rmse", n = 10)
(M1 <- select_best(hitters_tune, metric = "rmse"))
M2 <- select_by_one_std_err(hitters_tune, metric = "rmse", penalty)
select_by_pct_loss(hitters_tune, metric = "rmse", penalty, limit = 5)
```

## Finalise the Model

```{r}
hitters_final_model <- hitters_wf %>% finalize_workflow(M1)
hitters_final_model
```

## Variable Importance

```{r}
hitters_final_model %>%
  fit(hitters_train) %>%
  extract_fit_parsnip() %>%
  vip()
```

## Evaluate Model Performance Using Testing Data

```{r}
hitters_final_model %>% last_fit(hitters_split) %>% collect_metrics()
```

```{r}
hitters_test %>%
  add_column(hitters_final_model %>%
               fit(data = hitters_train) %>%
               predict(new_data = hitters_test)) %>%
  rmse(salary, .pred)
```


# DS_CS07_Decision_Trees_Titanic
## Libs

```{r}
pacman::p_load(tidyverse, tidymodels, janitor, visNetwork)
```

## Load data

```{r}
data("titanic_train", package = "titanic")
titanic <- as_tibble(titanic_train)
titanic
```

## Clean names

```{r}
titanic <- titanic %>% clean_names()
titanic
```

## Convert survived, pclass to factor

```{r}
titanic <- titanic %>% 
  mutate_at(vars(survived, pclass, sex), factor)
titanic
```

## Select survived, pclass, age, sib_sp, parch

```{r}
titanic <- titanic %>% select(survived, pclass, sex, age, sib_sp, parch)
titanic
```

## Set recipe

```{r}
titanc_recipe <- recipe(survived ~ ., data = titanic)
```

## Set model

```{r}
titanic_model <- decision_tree(mode = "classification") %>% 
  set_engine("rpart")
```

## Set workflow

```{r}
titanic_wf <- workflow() %>% 
  add_recipe(titanc_recipe) %>% 
  add_model(titanic_model)
```

## Fit model

```{r}
titanic_fit <- titanic_wf %>% fit(titanic) %>% pull_workflow_fit()
titanic_fit
```

```{r}
plot(titanic_fit$fit)
text(titanic_fit$fit)
```

```{r}
visTree(titanic_fit$fit)
```

## Predict

```{r}
jono <- tibble(
  pclass = "3", 
  sex = "male", 
  age = 51, 
  sib_sp = 1, 
  parch = 0
)
predict(titanic_fit, new_data = jono)
predict(titanic_fit, new_data = jono, type = "prob")
```


# DS_CS08_SVM_Board_Game
## Load data

```{r, eval = FALSE}
tuesdata <- tidytuesdayR::tt_load("2019-03-12")
board <- tuesdata$board_games
write_rds(board, glue::glue("{wdpath}/data/board.rds"))
```

```{r}
board <- read_rds(glue::glue("{wdpath}/data/board.rds"))
```

## Clean up

```{r}
board <- 
  board %>% 
  select(
    rating = average_rating, 
    time = playing_time, 
    year = year_published
  )
board
```

```{r}
board %>% ggplot(aes(time)) + geom_histogram(col = "black", fill = "orange")
```

```{r}
board <- 
  board %>% 
  filter(time > 0, time <= 360)
```

```{r}
board %>% ggplot(aes(time)) + geom_histogram(col = "black", fill = "orange")
```



Convert rating into three levels

```{r}
board <- 
  board %>% 
  mutate(
    rating = cut_number(
      rating, 3, 
      labels = c("pants", "yeah", "wow")
    )
  )
```


## EDA

```{r}
board %>% 
  ggplot(aes(year, time, 
             col = rating)) + 
  geom_point(alpha = 0.7) + 
  scale_colour_hp(house = "ravenclaw", discrete = TRUE)
  
```

```{r}
board %>% 
  ggplot(aes(year, time, 
             col = rating)) + 
  geom_density_2d() + 
  scale_colour_hp(house = "ravenclaw", discrete = TRUE)
```

## Model

```{r}
board_wf <- 
  workflow() %>% 
  add_recipe(
    recipe(rating ~ ., data = board) %>% 
      step_log(time)
  ) %>% 
  add_model(
    svm_poly(mode = "classification", degree = 2) %>% 
      set_engine("kernlab")
  )
```

```{r, cache=TRUE}
board_M1 <- board_wf %>% fit(board)
board_M1
```


```{r, cache=TRUE}
board_M2 <- 
  board_wf %>% 
  update_model(
    svm_rbf(mode = "classification") %>% 
      set_engine("kernlab")
  ) %>% 
  fit(board)
```


```{r, cache=TRUE}
board_M3 <- 
  board_wf %>% 
  update_model(
    rand_forest(mode = "classification") %>% 
      set_engine("ranger")
  ) %>% 
  fit(board)
```

## Predict

```{r}
newdata <- crossing(
  year = seq(1980, 2016, length = 100),
  time = seq(1, 360, length = 100)
)
```


```{r}
newdata %>% 
  add_column(
    predict(board_M1, new_data = newdata)
  ) %>% 
  ggplot(aes(year, time, fill = .pred_class)) + 
  geom_hex() + 
  scale_fill_hp(house = "ravenclaw", discrete = TRUE)
```


```{r}
newdata %>% 
  add_column(
    predict(board_M2, new_data = newdata)
  ) %>% 
  ggplot(aes(year, time, fill = .pred_class)) + 
  geom_hex() + 
  scale_fill_hp(house = "ravenclaw", discrete = TRUE)
```

```{r}
newdata %>% 
  add_column(
    predict(board_M3, new_data = newdata)
  ) %>% 
  ggplot(aes(year, time, fill = .pred_class)) + 
  geom_hex() + 
  scale_fill_hp(house = "ravenclaw", discrete = TRUE)
```


```{r}
board %>% 
  add_column(
    predict(board_M1, new_data = board, type = "prob")
  ) %>% 
  mutate(type = "SVM poly") %>% 
  bind_rows(
    board %>% 
      add_column(
        predict(board_M2, new_data = board, type = "prob")
      ) %>% 
      mutate(type = "SVM radial")
  ) %>% 
  bind_rows(
    board %>% 
      add_column(
        predict(board_M3, new_data = board, type = "prob")
      ) %>% 
      mutate(type = "RF")
  ) %>% 
  group_by(type) %>% 
  roc_curve(rating, starts_with(".pred")) %>% 
  autoplot()
```


# DS_CS09_PCA_Spotify
## Read in data

```{r}
spotify <- read_rds(glue::glue("{wdpath}/data/spotify.rds"))
spotify
```

## Clean up

```{r}
spotify <- 
  spotify %>% 
  select(danceability:duration_ms, genre = playlist_genre)
spotify
```

```{r, message=FALSE}
pacman::p_load(corrplot)
spotify %>% 
  select(danceability:duration_ms) %>% 
  cor(use = "pairwise.complete.obs") %>% 
  corrplot()
```

## PCA

```{r}
spotify_prep <- 
  spotify %>% 
  recipe(genre ~ .) %>% 
  step_normalize(all_predictors()) %>% 
  step_pca(all_predictors(), num_comp = 4) %>% 
  prep()
spotify_prep
```

## Loadings plot

```{r, message=FALSE}
pacman::p_load(tidytext)
spotify_prep %>% 
  tidy(n = 2) %>% 
  filter(component %in% c("PC1", "PC2", "PC3", "PC4")) %>% 
  group_by(component) %>% 
  top_n(5) %>% 
  mutate(
    terms_order = reorder_within(terms, abs(value), component)
  ) %>% 
  ggplot(
    aes(value, terms_order, fill = terms) 
  ) + 
  geom_col(show.legend = FALSE) + 
  facet_wrap(~component, scales = "free_y") + 
  scale_y_reordered()
```

## First two PC

```{r}
spotify_prep %>% 
  juice() %>% 
  ggplot(aes(PC1, PC2, col = genre))  + 
  geom_point(alpha = 0.7) + 
  scale_color_paletteer_d("calecopal::superbloom3")
```

```{r}
spotify_prep %>% 
  juice() %>% 
  ggplot(aes(PC1, fill = genre))  + 
  geom_density(alpha = 0.5) + 
  scale_fill_paletteer_d("calecopal::superbloom3")
```

## Skree plot

```{r}
sdev <- spotify_prep$steps[[2]]$res$sdev

percent_variation <- sdev^2 / sum(sdev^2)
percent_variation

tibble(
  component = str_c("PC", 1:length(percent_variation)),
  percent_var = percent_variation
) %>%
  mutate(component = fct_inorder(component)) %>%
  ggplot(aes(component, percent_var)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(x = NULL, y = "Percent variance explained by each PCA component")
```
