---
title: "STATS 7022 - Data Science <br>
 Seminar 12: Fitting Models (Solutions)"
author: "Max Glonek"
date: "Trimester `r if(as.integer(format(Sys.time(),'%m'))<5){1}else if(as.integer(format(Sys.time(),'%m'))<9){2}else{3}`, `r format(Sys.time(),'%Y')`"
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
    number_sections: true
    toc_depth: 3
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  fig.width = 6, 
  fig.asp = 0.618, 
  out.width = "70%",
  fig.align = "center", 
  root.dir = '../',
  warning=FALSE
)
pacman::p_load(
  bookdown, ggplot2
)
theme_set(theme_bw())
wdpath <- getwd()
```

# Load the Data and Libraries
```{r}
pacman::p_load(tidyverse, tidymodels, discrim, magrittr)
data <- read_rds("./data/data.rds")
```

#  Regression Models
## Linear Regression With Interaction Term
```{r}
# fit the model
lm_1 <- lm(X2 ~ X1*X3, data = data)

# fit with tidymodels
recipe_lm <- recipe(X2 ~ ., data = data) %>%
  step_interact(terms = ~ X1:X3)
recipe_lm %>% prep() %>% bake(new_data = NULL)
```


```{r}
model_lm <- linear_reg() %>% 
  set_engine("lm")
workflow_lm <- workflow() %>%
  add_recipe(recipe_lm) %>%
  add_model(model_lm)
fit_lm <- workflow_lm %>%
  fit(data)
fit_lm %>% tidy()
```

```{r}
# model summary
 summary(lm_1)
```

```{r}
# tidymodels summary
fit_lm %>% extract_fit_parsnip() %>% pluck("fit") %>% summary()
```

```{r}
metrics <- metric_set(rmse,rsq)
 (metrics_lm <- data %>% add_column(predict(fit_lm,data)) %>% metrics(truth = X2, estimate = .pred))
```

```{r}
 # visualise the data
 data %>%
 ggplot(aes(x = X1, y = X2, colour = X3)) +
 geom_point() +
 geom_smooth(method="lm")
```

```{r}
# check the assumptions
 plot(lm_1)
```

# Classification Models
##  Support Vector Machine With Gaussian Radial Basis Function Kernel

```{r}
# fit the model
svm_r <- recipe(X3 ~., data = data)
svm_m <- svm_rbf() %>%
  set_mode("classification") %>%
  set_engine("kernlab")
svm_wf <- workflow() %>%
  add_recipe(svm_r) %>%
  add_model(svm_m)
svm_fit <- svm_wf %>%
  fit(data)

# model summary
svm_fit %>% extract_fit_parsnip() %>% pluck("fit")
```

```{r}
# visualise the data
new_data <- expand.grid(
  X1 = seq(min(data$X1), max(data$X1), length.out = 200),
  X2 = seq(min(data$X2), max(data$X2), length.out = 200)
)

new_data  %>%
  add_column(predict(svm_fit, new_data = new_data)) %>%
  ggplot(aes(x=X1,y=X2,fill=.pred_class)) +
  geom_tile() +
  labs(fill = "X3",
       x = "X1",
       y = "X2") +
  theme_bw() +
  viridis::scale_fill_viridis(option="G",
                              begin = 1/6,
                              end = 5/6,
                              discrete=TRUE)
```


```{r}
# measure performance
svm_pred <- data %>%
  bind_cols(predict(svm_fit, data),
            predict(svm_fit, data, type = "prob"))
svm_pred %>% conf_mat(truth = X3,
                      estimate = .pred_class)

svm_pred %>% yardstick::roc_curve(truth = X3,
                                  .pred_0) %>%
  autoplot()
```

# LDA Without Interaction Term

```{r}
# fit the model
lda_r1 <- recipe(X3 ~ ., data = data)
lda_m1 <- discrim_linear() %>%
  set_mode("classification") %>%
  set_engine("MASS")
lda_wf1 <- workflow() %>%
  add_recipe(lda_r1) %>%
  add_model(lda_m1)
lda_fit1 <- lda_wf1 %>%
  fit(data)

# model summary
lda_fit1 %>% extract_fit_parsnip() %>% pluck("fit")
```

```{r}
# visualise the data
new_data %>%
  add_column(predict(lda_fit1, new_data = new_data)) %>%
  ggplot(aes(x=X1,y=X2,fill=.pred_class)) +
  geom_tile() +
  labs(fill = "X3",
       x = "X1",
       y = "X2") +
  theme_bw() +
  viridis::scale_fill_viridis(option="G",
                              begin = 1/6,
                              end = 5/6,
                              discrete=TRUE)
```

```{r}
# measure performance
lda_pred1 <- data %>%
  bind_cols(predict(lda_fit1, data),
            predict(lda_fit1, data, type = "prob"))
lda_pred1 %>% conf_mat(truth = X3,
                       estimate = .pred_class)

lda_pred1 %>% yardstick::roc_curve(truth = X3,
                                   .pred_0) %>%
  autoplot()
```

# LDA With Interaction Term

```{r}
# fit the model
lda_r2 <- recipe(X3 ~ ., data = data) %>%
  step_interact(terms = ~ X1:X2)
lda_m2 <- discrim_linear() %>%
  set_mode("classification") %>%
  set_engine("MASS")
lda_wf2 <- workflow() %>%
  add_recipe(lda_r2) %>%
  add_model(lda_m2)
lda_fit2 <- lda_wf2 %>%
  fit(data)

# model summary
lda_fit2 %>% extract_fit_parsnip() %>% pluck("fit")
```

```{r}
# visualise the data
new_data %>%
  add_column(predict(lda_fit2, new_data = new_data)) %>%
  ggplot(aes(x=X1,y=X2,fill=.pred_class)) +
  geom_tile() +
  labs(fill = "X3",
       x = "X1",
       y = "X2") +
  theme_bw() +
  viridis::scale_fill_viridis(option="G",
                              begin = 1/6,
                              end = 5/6,
                              discrete=TRUE)
```

```{r}
# measure performance
lda_pred2 <- data %>%
  bind_cols(predict(lda_fit2, data),
            predict(lda_fit2, data, type = "prob"))
lda_pred2 %>% conf_mat(truth = X3,
                       estimate = .pred_class)

lda_pred2 %>% yardstick::roc_curve(truth = X3,
                                   .pred_0) %>%
  autoplot()
```

# QDA
```{r}
# fit the model
qda_r <- recipe(X3 ~ ., data = data)
qda_m <- discrim_quad() %>%
  set_mode("classification") %>%
  set_engine("MASS")
qda_wf <- workflow() %>%
  add_recipe(qda_r) %>%
  add_model(qda_m)
qda_fit <- qda_wf %>%
  fit(data)

# model summary
qda_fit %>% extract_fit_parsnip() %>% pluck("fit")
```

```{r}
# visualise the data
new_data %>%
  add_column(predict(qda_fit, new_data = new_data)) %>%
  ggplot(aes(x=X1,y=X2,fill=.pred_class)) +
  geom_tile() +
  labs(fill = "X3",
       x = "X1",
       y = "X2") +
  theme_bw() +
  viridis::scale_fill_viridis(option="G",
                              begin = 1/6,
                              end = 5/6,
                              discrete=TRUE)
```

```{r}
# measure performance
qda_pred <- data %>%
  bind_cols(predict(qda_fit, data),
            predict(qda_fit, data, type = "prob"))
qda_pred %>% conf_mat(truth = X3,
                      estimate = .pred_class)

qda_pred %>% yardstick::roc_curve(truth = X3,
                                  .pred_0) %>%
  autoplot()
```

# Logistic Regression
```{r}
# fit the model
lr_r <- recipe(X3 ~ ., data = data)
lr_m <- logistic_reg() %>%
  set_mode("classification") %>%
  set_engine("glm")
lr_wf <- workflow() %>%
  add_recipe(lr_r) %>%
  add_model(lr_m)
lr_fit <- lr_wf %>%
  fit(data)

# model summary
lr_fit %>% extract_fit_parsnip() %>% pluck("fit")
```

```{r}
# visualise the data
new_data %>%
  add_column(predict(lr_fit, new_data = new_data)) %>%
  ggplot(aes(x=X1,y=X2,fill=.pred_class)) +
  geom_tile() +
  labs(fill = "X3",
       x = "X1",
       y = "X2") +
  theme_bw() +
  viridis::scale_fill_viridis(option="G",
                              begin = 1/6,
                              end = 5/6,
                              discrete=TRUE)
```

```{r}
# measure performance
lr_pred <- data %>%
  bind_cols(predict(lr_fit, data),
            predict(lr_fit, data, type = "prob"))
lr_pred %>% conf_mat(truth = X3,
                     estimate = .pred_class)

lr_pred %>% yardstick::roc_curve(truth = X3,
                                 .pred_0) %>%
  autoplot()
```

#  Comparing Model Performance
```{r}
# get predictions
preds <- rbind(lda_pred1 %>% add_column(model = "LDA +"),
               lda_pred2 %>% add_column(model = "LDA x"),
               qda_pred %>% add_column(model = "QDA"),
               svm_pred %>% add_column(model = "SVM (RBF)"),
               lr_pred %>% add_column(model = "Logsitic Reg"))


# plot ROC curves
roc_df <- preds %>%
  group_by(model) %>%
  group_map(~ yardstick::roc_curve(.x, truth = X3, .pred_0) %>% mutate(model = .y$model)) %>%
  bind_rows()

ggplot(roc_df, aes(x = 1 - specificity, y = sensitivity, color = model)) +
  geom_path(size = 1) +
  viridis::scale_color_viridis(option = "D", begin = 1/6, end = 5/6, discrete = TRUE) +
  geom_abline(lty = 2) +
  coord_equal() +
  theme_bw()
```

```{r}
# auc
preds %>%
  group_by(model) %>%
  group_map(~ yardstick::roc_auc(.x, truth = X3, .pred_0) %>%
              mutate(model = .y$model)) %>%
  bind_rows()
```

#  Which Model Should We Choose?
 The LDA model without an interaction term is clearly rubbish, as is the logistic regression model. We should notuse either of these models. It might be worth fitting a logistic regression model with an interaction term - I
 suspect it would perform quite well.
```{r}
# assumption checking for LDA/QDA

## group 0
### X1 qqplot
qqnorm(data$X1[data$X3==0])
qqline(data$X1[data$X3==0])
```

```{r}
### X2 qqplot
qqnorm(data$X2[data$X3==0])
qqline(data$X2[data$X3==0])
```

```{r}
### variances and correlation
var(data$X1[data$X3==0])

var(data$X2[data$X3==0])

cor(data$X1[data$X3==0],data$X2[data$X3==0])
```

```{r}
## group 1
### X1 qqplot
qqnorm(data$X1[data$X3==1])
qqline(data$X1[data$X3==1])
```

```{r}
### X2 qqplot
qqnorm(data$X2[data$X3==1])
qqline(data$X2[data$X3==1])
```

```{r}
### variances and correlation
var(data$X1[data$X3==1])

var(data$X2[data$X3==1])

cor(data$X1[data$X3==1],data$X2[data$X3==1])
```

 LDA assumes multivariate normal for each group - qqplots show this assumption is reasonable. LDA also assumes constant variance/covariance structure for all groups - comparing $cor(X_1, X_2)$ for each group shows this assumption is unreasonable. Thus, we should not use LDA.

 QDA relaxes the assumption of constant variance/covariance structure, so would be a reasonable model to choose in this instance. SVM would also be a reasonable model, though its performance is marginally worse than QDA, and we may want to choose our kernel more carefully


 
