---
title: 'Linear Regression: Fuel Economy'
author: "Max Glonek"
date: Trimester `r if(as.integer(format(Sys.time(),'%m'))<5){1}else if(as.integer(format(Sys.time(),'%m'))<9){2}else{3}`,
  `r format(Sys.time(),'%Y')`
output:
  pdf_document:
    toc: true
    latex_engine: xelatex
  html_document:
    number_sections: true
    df_print: paged
    toc: true
    toc_float: true
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  fig.width = 6, 
  fig.asp = 0.618, 
  out.width = "70%",
  fig.align = "center", 
  root.dir = '../'
)
```

# Case study

## Set up libraries

```{r}
pacman::p_load(
  tidyverse, 
  tidymodels,
  AppliedPredictiveModeling, 
  skimr, 
  janitor, 
  corrplot, 
  vip
)
```

## Read in data

```{r}
data(FuelEconomy, package = "AppliedPredictiveModeling")
```


## Pre-processing

### Convert to tibble

```{r}
cars2010 <- as_tibble(cars2010)
```

### Clean names and select predictors

```{r}
cars2010 <- clean_names(cars2010)
cars2010 <- 
  cars2010 %>% 
  select(
    eng_displ, 
    num_cyl, 
    FE = fe, 
    num_gears, 
    drive_desc
  )
cars2010
```

### Recode the drive predictor

```{r}
cars2010 %>% 
  count(drive_desc)
cars2010<- 
  cars2010 %>% 
  mutate(
    drive = case_when(
      str_detect(drive_desc, "Front") ~ "front", 
      str_detect(drive_desc, "Rear") ~ "rear", 
      TRUE ~ "4WD"
    ))
cars2010 %>% 
  count(drive_desc, drive)
cars2010 <- 
  cars2010 %>% 
  select(-drive_desc)
cars2010
```

### Convert character into factor

```{r}
cars2010 <- 
  cars2010 %>% 
  mutate_if(
    is.character, factor
  )
cars2010
```


### Split into training and testing

```{r}
set.seed(2020)
car_split <- initial_split(cars2010, strata = FE)
car_train <- training(car_split)
car_test <- testing(car_split)
```


## EDA

### Overview

```{r}
skim(car_train)
```


### Look at Fuel efficiency

```{r}
car_train %>% 
  ggplot(aes(FE)) + 
  geom_histogram(col = "black", fill = "orange")

```

### Bivariate relationships

```{r}
car_train %>% 
  ggplot(aes(eng_displ, FE)) + 
  geom_point()
```

```{r}
car_train %>% 
  ggplot(aes(factor(num_cyl), FE)) + 
  geom_boxplot()
```

```{r}
car_train %>% 
  ggplot(aes(factor(num_gears), FE)) + 
  geom_boxplot()
```

```{r}
car_train %>% 
  ggplot(aes(drive, FE)) + geom_boxplot()
```

### Relationship between predictors

```{r}
car_train %>% 
  select_if(is.numeric) %>% 
  cor() %>% 
  corrplot()
```


### Preprocessing with recipes

```{r}
car_recipe <- 
  recipe(FE ~ ., data = car_train) %>% 
  step_poly(eng_displ, degree = 2, options = list(raw = TRUE))
car_recipe
```


```{r}
car_recipe <- car_recipe %>% prep()
car_recipe
```


```{r}
car_train <- car_recipe %>% juice()
car_train
```


```{r}
car_test <- car_recipe %>% bake(car_test)
car_test
```

## Fit model

```{r}
car_lm <- 
  linear_reg() %>% 
  set_engine("lm") %>% 
  fit(FE ~ ., data = car_train)
car_lm
```


## Use model

### Model fit

```{r}
car_train %>% 
  add_column(
    predict(car_lm, new_data = car_train)
  ) %>% 
  yardstick::rmse(truth = FE, estimate = .pred)
```

```{r}
car_test %>% 
  add_column(
    predict(car_lm, new_data = car_test)
  ) %>% 
  yardstick::rmse(truth = FE, estimate = .pred)

```

### Check assumptions

```{r}
car_res <- augment(car_lm$fit)
car_res
```

```{r}
car_res %>% 
  ggplot(aes(.fitted, .resid)) + 
  geom_point() + 
  geom_hline(yintercept = 0)
```

```{r}
car_res %>% 
  ggplot(aes(sample = .resid)) + 
  stat_qq() + 
  stat_qq_line()
```

```{r}
car_res %>% 
  ggplot(aes(.hat, .std.resid)) +
  geom_vline(size = 2, colour = "white", xintercept = 0) +
  geom_hline(size = 2, colour = "white", yintercept = 0) +
  geom_point(aes(size = .cooksd)) + geom_smooth(se = FALSE)
```

### Variable importance

```{r}
vi(car_lm)
```

```{r}
vip(car_lm, 
    include_type = TRUE, 
    geom = 'point',
    mapping = aes(col = "Sign"), size = 5) + 
  scale_color_brewer(palette = "Set1")
```

### Inference

```{r}
tidy(car_lm$fit)
```

```{r}
tidy(car_lm$fit) %>% 
  arrange(p.value)
```

### Prediction


```{r}
new_data <- tibble(
  num_cyl = 4, 
  num_gears = 6, 
  eng_displ = 5, 
  drive = "front"
) 
new_data
```

```{r}
new_data <- bake(car_recipe, new_data = new_data)
new_data
```

```{r}
predict(car_lm, new_data = new_data)
predict(car_lm, new_data = new_data, type = "pred_int")
```

