---
title: |
    | STATS 7022 - Data Science
    | Workshop 6: Nested Data Frames
author: "Max Glonek"
institute: "School of Computer and Mathematical Sciences, University of Adelaide"
date: "Trimester `r if(as.integer(format(Sys.time(),'%m'))<5){1}else if(as.integer(format(Sys.time(),'%m'))<9){2}else{3}`, `r format(Sys.time(),'%Y')`"
output:
  html_document: default
  beamer_presentation: 
    includes: 
      in_header: header.tex
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "100%",
  fig.align = "center",
  root.dir = '../'
)
```

## Set-Up
```{r}
pacman::p_load(tidyverse)
```

## Lists
In `R`, elements of a vector must all be the same type, e.g., numeric, integer, boolean, character.

By contrast, a list is a collection of objects of *any* type.

We have been using lists all along. `data.frame` and `tibble` are both examples of lists (with some cleverness applied under the hood).

Lists behave much like other objects in `R`, but are indexed with `[[ ]]`.

## Map Functions
A `map()` function takes a list as input, does *something*, and returns a list.

```{r}
lst <- list(1:10, c(2, 4))
lst %>% map(mean)
```

## Dialects of Map
Sometimes you don't want to return a list, e.g., return results as a vector with `double` precision.

```{r}
lst %>% map_dbl(mean)
```

## Fast Functions
```{r}
lst %>% map(~.x + 3)
```

## Data
\small
```{r}
pacman::p_load(gapminder)
data(gapminder)
gapminder
```

## Life Expectancy
```{r, echo=FALSE}
gapminder %>%
  ggplot(aes(year, lifeExp, group = country)) +
  geom_line(alpha = 1/5)
```

## Nest by Country
\scriptsize
```{r}
gapminder_nest <- gapminder %>%
  group_by(country, continent) %>%
  nest()
gapminder_nest
```

## Getting the Data Back for Each Country
\small
```{r}
gapminder_nest$data[[1]]
```

## A Regression Function
Here's a function to fit a linear regression of life expectancy on year for data contained within *some* data.frame.

```{r}
fit_lm <- function(df){
  lm(lifeExp ~ year, data = df)
}
```

## Lots of Linear Models
```{r}
gapminder_nest <- gapminder_nest %>%
  mutate(model = map(data,
                     fit_lm))
```

## Getting the Model Back for Each Country
```{r}
gapminder_nest$model[[1]]
```

## Getting the Residuals
```{r}
gapminder_nest <- gapminder_nest %>%
  mutate(residuals = map(model,
                         broom::augment))
```

## Unnesting the Data Frame
\tiny
```{r}
gapminder_nest %>%
  unnest(residuals)
```

## Task
```{r, echo=TRUE}
gapminder %>% 
  mutate(
    year = year - mean(year)
  ) %>% 
  group_by(
    country, continent
  ) %>% 
  nest() %>% 
  mutate(
    model = map(data, fit_lm), 
    coef = map(model, broom::tidy)
  ) %>% 
  unnest(coef) %>% 
  select(country, continent, term, estimate) %>% 
  pivot_wider(names_from = term, values_from = estimate) %>% 
  rename(intercept = `(Intercept)`, slope = year) %>% 
  ggplot(aes(intercept, slope, col = continent, shape = continent)) + 
  geom_point() + 
  viridis::scale_colour_viridis(option="viridis", discrete=TRUE)
```