---
title: "STATS 7022 - Data Science <br> 
Assignment 2: Question 1"
author: "Dongju Ma"
institute: "School of Computer and Mathematical Sciences, University of Adelaide"
date: "Trimester `r if(as.integer(format(Sys.time(),'%m'))<5){1}else if(as.integer(format(Sys.time(),'%m'))<9){2}else{3}`, `r format(Sys.time(),'%Y')`"
output:
  bookdown::html_document2: default
  bookdown::pdf_document2:
    toc: no
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
 echo = TRUE, 
 fig.width = 6, 
 fig.asp = 0.707, 
 out.width = "100%",
 fig.align = "center", 
  root.dir = '../'
)

pacman::p_load(bookdown, kableExtra)

options(knitr.table.format = "html")  
 
```

# Load Libraries

```{r LoadLibraries}
pacman::p_load(dplyr, tidymodels)
```

# Load Data

The data is contained within the file `bird_bath.rds` . The data is read in to R .  

```{r LoadData}
bird_bath <- readRDS("./data/bird_bath.rds")
bird_bath
```

There are 161,057 rows in the `bird_bath` data.

# Data Cleanning

## Remove Missing Years

Some rows are missing a survey year ( `survey_year` ). These rows are removed from the data.  

```{r RemoveNA}
bird_bath <- bird_bath %>% na.omit(survey_year)
bird_bath
```

There are now 160,888 rows in the `bird_bath` data.

## Identify the Top 15 Types of Bird

We consider only the top 15 bird species ( `bird_type` ) by frequency ( `bird_count` ). Results are summarised in Table \@ref(tab:Top15).

```{r Top15}
top_15_birds <- filter(bird_bath, bird_count == 1)%>%
  count(bird_type, sort = TRUE) %>%
  slice_head(n = 15)

top_15_birds %>% 
  knitr::kable(caption = "Top 15 birds at Australian bird baths, by frequency.") %>% 
  kable_styling(full_width = TRUE)
```

```{r Top15Filter}
bird_bath <- bird_bath %>%
  filter(bird_type %in% top_15_birds$bird_type)
bird_bath
```

There are now 14,280 rows in the `bird_bath` data.
 
## Create the Response Variable

We define a new response variable, called `bird_present` , using the following rule:

- If `bird_count` is greater than zero for a row, then `bird_present` takes value “yes”,  
- Otherwise, `bird_present` takes value “no”.

```{r ResponseVariable}
bird_bath <- bird_bath %>% 
  mutate(bird_present = ifelse(bird_count > 0, "yes", "no"))
bird_bath
```

## Select Columns

We only need the following columns:

 - `urban_rural`  
 - `bird_type`  
 - `bird_present`

```{r Selection} 
bird_bath <- bird_bath %>% 
  select(urban_rural, bird_type, bird_present)
bird_bath
```

All other columns are removed.

##  Convert Variables to Categorical

We convert all remaining variables to categorical variables.

```{r Factors}
bird_bath <- bird_bath %>% 
  mutate_all(factor)
bird_bath
```

# Models

We create two separate workflows using `tidymodels` :  

 1. A logistic regression model with `bird_present` as the response variable and urban_rural and bird_type
 as the predictors

```{r WF1}
bird_bath_recipe1 <- recipe(bird_present ~ urban_rural + bird_type, data = bird_bath) %>%
  step_dummy()

bird_bath_model1 <- logistic_reg() %>% 
  set_mode("classification") %>% 
  set_engine("glm")

bird_bath_wf1 <- workflow() %>% 
  add_recipe(bird_bath_recipe1) %>% 
  add_model(bird_bath_model1)
bird_bath_wf1
```

 2. A logistic regression model with the same predictors and response, plus a two-way interaction between
 `urban_rural` and bird_type 
 
```{r WF2}
bird_bath_recipe2 <- recipe(bird_present ~ urban_rural + bird_type, data = bird_bath) %>%
  step_dummy() %>% 
  step_interact(bird_present ~ urban_rural:bird_type)

bird_bath_model2 <- logistic_reg() %>% 
  set_mode("classification") %>% 
  set_engine("glm")

bird_bath_wf2 <- workflow() %>% 
  add_recipe(bird_bath_recipe2) %>% 
  add_model(bird_bath_model2)
bird_bath_wf2
```
Both models are fitted to the `bird_bath` data, with ROC curves displayed in Figure \@ref(fig:ROC)

# ROC Curves

```{r ROC, fig.cap="ROC curves for logisitic regression models."}
bird_bath_fit1 <- bird_bath_wf1 %>% 
  fit(bird_bath)

bird_bath_fit2 <- bird_bath_wf2 %>% 
  fit(bird_bath)

bird_bath_pred1 <- bird_bath %>%
  bind_cols(predict(bird_bath_fit1, bird_bath),
            predict(bird_bath_fit1, bird_bath, type = "prob")) %>% 
  select(bird_present, starts_with(".pred"))

bird_bath_pred2 <- bird_bath %>%
  bind_cols(predict(bird_bath_fit2, bird_bath),
            predict(bird_bath_fit2, bird_bath, type = "prob")) %>%
  select(bird_present, starts_with(".pred"))

roc_1 <- bird_bath_pred1 %>% 
  roc_curve(truth = bird_present, .pred_no) %>% 
  mutate(model = "Model 1")

roc_2 <-  bird_bath_pred2 %>% 
  roc_curve(truth = bird_present, .pred_no) %>% 
  mutate(model = "Model 2")

roc_bind <- bind_rows(roc_1, roc_2) %>%
  mutate(model = factor(model, levels = c("Model 1", "Model 2")))
roc_bind
 
ggplot(roc_bind, aes(x = 1 - specificity, y = sensitivity, color = model)) +
  geom_line(linewidth = 0.5) +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted") +
    scale_color_manual(values = c("Model 1" = "#F66359", "Model 2" = "#1AC5CA")) +
  labs(color = "model") +
  coord_equal() +
  theme_bw() +
  theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.position = "right"
  )
```

