---
title: "STATS 7022 - Data Science Assignment 3: Question 1"
author: "Dongju Ma"
institute: "School of Computer and Mathematical Sciences, University of Adelaide"
date: "Trimester `r if(as.integer(format(Sys.time(),'%m'))<5){1}else if(as.integer(format(Sys.time(),'%m'))<9){2}else{3}`, `r format(Sys.time(),'%Y')`"
output:
  # bookdown::pdf_document2:
  #   toc: no
  #   latex_engine: xelatex
  #   number_sections: false
  bookdown::html_document2: 
    number_sections: false
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(
 echo = TRUE, 
 fig.width = 6, 
 fig.asp = 0.618, 
 out.width = "70%",
 fig.align = "center", 
  root.dir = '../'
)

pacman::p_load(bookdown, kableExtra)
options(knitr.table.format = "html")
```

# (a) Read in the data
```{r data}
pacman::p_load(tidyverse, tidymodels, kernlab)
a3 <- read_rds('./data/a3_data.rds')
a3
```
There are 6,000 observations in the original data.


# (b) EDA
```{r skim1}
skimr::skim(a3) 
```

# (c) Clear the data
## i. Data types
As the output in Table \@ref(tab:skim1), the data type of X1 is character and which should be float.  
The data type of X1 could be corrected by the following code.
```{r}
a3 <- a3 %>% 
  mutate(X1 = as.numeric(X1))
```

## ii. Removing numerical outliers
With the numerical part of Table \@ref(tab:skim1), the P0 of X2 and the P100 of X4 shows there are outliers in these two variables, which could be filtered by the following R code,
```{r filter_num}
a3 %>% 
  filter(X2 < -20)
a3 %>% 
  filter(X4 > 20)
```

and then replace them by NA.
```{r replace_num}
a3 <- a3 %>% 
  mutate(X2 = case_when(X2 >= -20 ~ X2,
                        TRUE ~ NA)) %>% 
  mutate(X4 = case_when(X4 <= 20 ~ X4, 
                        TRUE ~ NA))
```

## iii. Removing categorical outliers
As Table \@ref(tab:skim1) shows, Y has 4 levels which should be only 2(A or B), while C2 has 4 levels which should be 3(V, U, T). The following work is to clean the unexpected levels.
```{r replace_cate}
a3 <- a3 %>% 
  mutate(Y = case_when(Y %in% c("A", "B") ~ Y,
                       TRUE ~ NA)) %>% 
  mutate(C2 = case_when(C2 %in% c("T", "U", "V") ~ C2,
                        TRUE ~ NA)) %>% 
  mutate(Y = factor(Y, levels = c("A", "B"))) %>% 
  mutate(C2 = factor(C2, levels = c("T", "U", "V")))
```

## iv. Removing NA
```{r naomit}
a3 <- a3 %>% 
  na.omit() 
```

## v. Final check
```{r skim2}
skimr::skim(a3)
```
All the variables are normal and correct, without any missing values, there are now 5614 observations left.

# (d) Recipe
```{r recipe}
a3_recipe <- recipe(Y ~ ., data = a3) %>%
  step_normalize(all_numeric_predictors()) %>% 
  step_dummy(all_nominal_predictors())
a3_recipe
```

# (e) SVM models
```{r models}
m1 <- svm_poly(mode = "classification", degree = 3) %>% 
  set_engine("kernlab") 
  
m2 <- svm_rbf(mode = "classification") %>% 
  set_engine("kernlab")
```

# (f) Workflows
```{r wf}
a3_wf1 <- workflow() %>% 
  add_recipe(a3_recipe) %>% 
  add_model(m1)

a3_wf2 <- workflow() %>% 
  add_recipe(a3_recipe) %>% 
  add_model(m2)
```

# (g) Models Fitting
```{r fit}
a3_poly <- a3_wf1 %>% 
  fit(a3)
a3_poly

a3_radial <- a3_wf2 %>% 
  fit(a3)
a3_radial

a3_pred <- a3 %>% 
  bind_cols(
    predict(a3_poly, a3),
    predict(a3_poly, a3, type = "prob")
    ) %>% 
  mutate(type = "SVM Poly") %>% 
  bind_rows(
    a3 %>% 
      bind_cols(
        predict(a3_radial, a3),
        predict(a3_radial, a3, type = "prob")
        ) %>% 
      mutate(type = "SVM Radial")
  )
a3_pred

```

# (h) Performance metrics
## i. Accuracy
```{r acc}
acc <- a3_pred %>% 
  group_by(type) %>% 
  accuracy(
    truth = Y, estimate = .pred_class)
acc

```

## ii. Sensitivity
```{r sens}
sens <- a3_pred %>% 
  group_by(type) %>% 
  sens(truth = Y, estimate = .pred_class)
sens

```

## iii. Specificity
```{r spec}
spec <- a3_pred %>% 
  group_by(type) %>% 
  spec(truth = Y, estimate = .pred_class)
spec

```

# (i) ROC curve and ROC-AUC
```{r rocplot, fig.cap = "ROC Curves of SVM models." } 
a3_pred %>% 
  group_by(type) %>% 
  roc_curve(Y,
            .pred_A) %>% 
  autoplot()

```
The AUC of curves are
```{r auc}
auc <- a3_pred %>% 
  group_by(type) %>% 
  roc_auc(Y, .pred_A)
auc

```

# (j) Select the model
```{r metrics, echo=FALSE}
a3_metrics <- acc %>%
  select(type, .metric, .estimate) %>%
  bind_rows(
    sens %>% select(type, .metric, .estimate),
    spec %>% select(type, .metric, .estimate),
    auc %>% select(type, .metric, .estimate)
  ) %>%
  pivot_wider(names_from = .metric, values_from = .estimate) 
a3_metrics %>% 
  knitr::kable(caption = "Metrics Table for SVM models.") %>% 
  kable_styling(full_width = TRUE)
```
As Table \@ref(tab:metrics) shows, the SVM model with a polynomial kernel performance is better.

# (k) Prediction
```{r newdata, echo=FALSE}
new_data <- data.frame(
  "X1" = -0.11,
  "X2" = 1.19,
  "X3" = -1.37,
  "X4" = -0.46,
  "C1" = "y",
  "C2" = "T"
)

new_data_long <- tibble::enframe(as.list(new_data), name = "Predictor", value = "Value") 

new_data_long %>% 
  knitr::kable(caption = "New data for Prediction.") %>% 
  kable_styling(full_width = TRUE)
```
Use the SVM model with a polynomial kernel to make prediction of the data in Table \@ref(tab:newdata).
```{r prediction}
new_data <- data.frame(
  "X1" = -0.11,
  "X2" = 1.19,
  "X3" = -1.37,
  "X4" = -0.46,
  "C1" = "y",
  "C2" = "T"
)
predict(a3_poly, new_data)
```
And the result is A.
