---
title: "DS_A1_Q2"
author: "Dongju Ma"
institute: "School of Computer and Mathematical Sciences, University of Adelaide"
date: "Trimester `r if(as.integer(format(Sys.time(),'%m'))<5){1}else if(as.integer(format(Sys.time(),'%m'))<9){2}else{3}`, `r format(Sys.time(),'%Y')`"
output:
  bookdown::pdf_document2:
    toc: no
    includes:
      in_header: header.tex
  bookdown::html_document2: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
 echo = TRUE, 
 fig.width = 6, 
 fig.asp = 0.618, 
 out.width = "70%",
 fig.align = "center", 
  root.dir = '../'
)
pacman::p_load(tidyverse,tidymodels,bookdown)
```
# Q2. (a) Fair coin
Since we are using the fair coin with tossed to the outcome of half heads and tails, we can easily get the confusion matrices with $n$ times of experiments like Table \@ref(tab:cmfcTab), as we make an assumption that heads are true, and the observed proportion of truth is $p, (0 \leq p \leq 1)$.
```{r cmfcTab, echo= FALSE}
conf_matrix <- matrix(
  c("$\\frac{np}{2}$", 
    "$\\frac{n(1-p)}{2}$", 
    "$\\frac{np}{2}$", 
    "$\\frac{n(1-p)}{2}$" ), 
  nrow = 2, byrow = TRUE)
rownames(conf_matrix) <- c("Predicted Positive", "Predicted Negative")
colnames(conf_matrix) <- c("Truth Positive", "Truth Negative")
knitr::kable(conf_matrix, caption = "Confusion Matrix of Fair Coin.", escape = FALSE)
```

## (i) The expected sensitivity
So we could calculate the expected sensitivity using below formula,
$$
\text{Sensitivity} = \frac{TP}{TP + FN} = \frac{\frac{np}{2}}{\frac{np}{2} + \frac{np}{2}} = \frac{1}{2}
$$

## (ii) The expected specificity
Also we could calculate the expected specificity by
$$
\text{Specificity} = \frac{TN}{TN + FP} = \frac{\frac{n(1 - p)}{2}}{\frac{n(1-p)}{2} + \frac{n(1-p)}{2}} = \frac{1}{2}
$$

# Q2. (b) Biased Coin
Still we should figure out the confusion matrices first, as the biased coin have a probability of $q, (0 \leq q \leq 1)$ to be tossed with heads outcomes. This time we still have observed proportion $p,(0 \leq q \leq 1)$ of heads, so the matrices should be like Table \@ref(tab:cmbcTab).
```{r cmbcTab, echo = FALSE}
conf_matrix <- matrix(
  c("$nqp$", 
    "$nq(1-p)$", 
    "$np(1-q)$", 
    "$n(1-q)(1-p)$" ), 
  nrow = 2, byrow = TRUE)
rownames(conf_matrix) <- c("Predicted Positive", "Predicted Negative")
colnames(conf_matrix) <- c("Truth Positive", "Truth Negative")
knitr::kable(conf_matrix, caption = "Confusion Matrix of Biased Coin.", escape = FALSE)
```
This time we could calculate the biased sensitivity and specificity by
$$
\begin{aligned}
\text{Sensitivity} &= \frac{TP}{TP + FN} = \frac{nqp}{nqp + np(1-q)} = \frac{nqp}{np} = q \\
\text{Specificity} &= \frac{TN}{TN + FP} = \frac{n(1-q)(1-p)}{nq(1-p) + n(1-q)(1-p)} = \frac{n(1-q)(1-p)}{n(1-p)} = 1-q
\end{aligned}
$$

# Q2. (c) Conclusions
We could make following conclusions from these results,

* If the truth condition happens with probability $P, (0 \leq P \leq 1)$, the sensitivity would be $P$ while the specificity should be $1 - P$.  
* The sum of sensitivity and specificity should be 1.  


