---
title: "DS_A1_Q1"
author: "Dongju Ma"
institute: "School of Computer and Mathematical Sciences, University of Adelaide"
date: "Trimester `r if(as.integer(format(Sys.time(),'%m'))<5){1}else if(as.integer(format(Sys.time(),'%m'))<9){2}else{3}`, `r format(Sys.time(),'%Y')`"
output:
  bookdown::pdf_document2:
    toc: no
  bookdown::html_document2: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
 echo = TRUE, 
 fig.width = 6, 
 fig.asp = 0.618, 
 out.width = "70%",
 fig.align = "center", 
  root.dir = '../'
)
pacman::p_load(tidyverse,tidymodels,bookdown)
source("q1.R")
```

# R code for creating column plot of mean fuel efficiency
To create a bar chart for average fuel efficiencies of city and highway grouped by car classes, we should group the data and calculate their each mean fuel efficiencies, as Table \@ref(tab:meaneffTab) shows.
```{r code=readLines("q1.R"), eval=FALSE}
```
```{r meaneffTab, echo=FALSE}
mean_eff %>% 
  knitr::kable(caption = "Mean Fuel Efficiencies grouped by Vehicle Classes.")
```

An important aspect to highlight is that the bars in the plot should be ordered from lowest to highest based on the city's mean fuel efficiency. Therefore, we need to use R code to ensure the bars are plotted in the correct order.    

# Generate the plot
- **The figure should have titles**  
- **The colours filled in the bars**: #0d0887 for city and #e16462 for highway  
- **The bars of highways** should be placed next to those of cities  
- **The legend** should be placed at the bottom of the figure  

```{r Plot, fig.cap="Column plot of mean fuel efficiency", echo=FALSE}
# Plot
ggplot(mean_eff, aes(x = class, y = mean, fill = measure)) +
  geom_col(position = "dodge", color = "black", linewidth = 0.5) +
  # Add title and axis title
  labs(
    title = "Column plot of mean fuel efficiency",
    x = "Vehicle class",
    y = "Mean fuel efficiency (miles per gallon)"
  ) +
  # Set filling colours and themes
  scale_fill_manual(values = c("cty" = "#0d0887", "hwy" = "#e16462")) +
  theme_bw() +
  theme(legend.position = "bottom") 
```

So we would generate the Figure \@ref(fig:Plot) as the question asked. 