---
title: "Ridge Regression: The Office"
author: "Max Glonek"
date: "Trimester `r if(as.integer(format(Sys.time(),'%m'))<5){1}else if(as.integer(format(Sys.time(),'%m'))<9){2}else{3}`, `r format(Sys.time(),'%Y')`"
output:
  html_document:
    number_sections: yes
    df_print: paged
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
 echo = TRUE, 
 fig.width = 6, 
 fig.asp = 0.618, 
 out.width = "70%",
 fig.align = "center", 
  root.dir = '../'
)
pacman::p_load(tidyverse, tidymodels, tidytuesdayR, schrute, glmnet, plotly,
               vip, paletteer)
theme_set(theme_bw())
```

# Read in data

```{r}
data("theoffice", package = "schrute")
office <- theoffice
office
```

So each row is a line from the script, we also have info about season, episode etc. 

Want to predict rating from how much each character speaks. 

# Characters

How many characters?

```{r}
length(unique(office$character))
```

Too many, so we will collapse down

```{r}
office <- 
  office %>% 
  mutate(
    character = fct_lump_n(character, 20)
  )
unique(office$character)
```

# Cleaning

First count the number of lines for each character. 

```{r}
office_long <- 
  office %>% 
  group_by(season, episode, character) %>% 
  summarise(
    n_lines = n(), 
    rating = mean(imdb_rating)
  )
office_long
```

Now convert to wide form

```{r}
office_wide <- 
  office_long %>% 
  pivot_wider(
    names_from = character, 
    values_from = n_lines, 
    values_fill = 0
  )
office_wide
```

# Split

```{r}
office_split <- initial_split(office_wide, strata = rating)
office_train <- training(office_split)
office_test <- testing(office_split)
```

# Preprocessing

```{r}
office_recipe <- 
  recipe(rating ~ ., data = office_train) 
office_recipe
```

Set ID columns

```{r}
office_recipe <- 
  office_recipe %>% 
  update_role(season,episode, new_role = "ID")
office_recipe
```

Scale and center

```{r}
office_recipe <- 
  office_recipe %>% 
  step_center(all_predictors()) %>% 
  step_scale(all_predictors())
```

Check 

```{r}
office_recipe %>% prep() %>% juice()
office_recipe %>% prep() %>% juice() %>% select(rating)
```

# Fit ridge regression

## Set model

```{r}
office_model <- 
  linear_reg(mixture = 0, penalty = 0.1) %>% 
  set_engine("glmnet")
```

Note still uses lots of lambdas - see fit. Penalty only used for prediction. 

## Create workflow

```{r}
office_wf <- 
  workflow() %>% 
  add_recipe(office_recipe) %>% 
  add_model(office_model)
office_wf
```

## Fit workflow

```{r}
office_ridge <- 
  office_wf %>% 
  fit(office_train)
office_ridge
```

```{r}
p <- office_ridge %>% 
  extract_fit_parsnip() %>% 
  .$fit  %>% 
  tidy() %>% 
  filter(
    term != "(Intercept)", 
    lambda < 100
  ) %>% 
  ggplot(aes(lambda, estimate, col = term)) + 
  geom_line() 
p
```

```{r}
ggplotly(p)
```

# Predict

## Train RMSE

```{r}
office_train %>% 
  add_column(
    predict(office_ridge, new_data = office_train)
  ) %>% 
  ungroup() %>% 
  rmse(rating, .pred)
```

## Test RMSE

```{r}
office_test %>% 
  add_column(
    predict(office_ridge, new_data = office_test)
  ) %>% 
  ungroup() %>% 
  rmse(rating, .pred)
```

# VIP

```{r}
office_ridge %>% 
  pull_workflow_fit() %>% 
  vi() %>% 
  ggplot(aes(Importance, fct_reorder(Variable, Importance), fill = Sign)) + 
  geom_col() + 
  labs(y = "Character") + 
  scale_fill_paletteer_d("dutchmasters::milkmaid")
```

```{r}
sessionInfo()
```