---
title: "Seminar 3: Solutions"
author: "Max Glonek"
institute: "School of Computer and Mathematical Sciences, University of Adelaide"
date: "Trimester `r if(as.integer(format(Sys.time(),'%m'))<5){1}else if(as.integer(format(Sys.time(),'%m'))<9){2}else{3}`, `r format(Sys.time(),'%Y')`"
output:
  bookdown::pdf_document2:
    toc: no
    includes:
      in_header: header_sol.tex
  bookdown::html_document2: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
 echo = TRUE, 
 fig.width = 6, 
 fig.asp = 0.618, 
 out.width = "70%",
 fig.align = "center", 
  root.dir = '../'
)
pacman::p_load(tidyverse,tidymodels,bookdown)
```

# Try it Yourself: Add a Figure

```{r longleyPlot, fig.cap="Scatterplot of Employed vs GNP for the longley dataset", echo=FALSE}
data(longley)
plot(Employed ~ GNP, data=longley, col="#3fbf7f", pch=20, cex=2)
abline(lm(Employed ~ GNP, data=longley),col="#bf7fff",lwd=2)
```

Figure \@ref(fig:longleyPlot) shows a strong, positive, linear relationship between number of employed people and Gross National Product.

# Try it Yourself: Add a Table

```{r toothTab, echo=FALSE}
data(ToothGrowth)
ToothGrowth %>%
  count(supp) %>%
  knitr::kable(caption = "ToothGrowth factors.")
```

Table \@ref(tab:toothTab) shows a balanced study design in the ToothGrowth dataset.

# Try it Yourself: Formatting Mathematics

For a discrete random variable, $X$, the expected value of \(X\) is calculated as:

$$
E[X] = \frac{1}{n}\sum_{i=1}^{n}X_{i}
$$

We can also calculate the variance of \(X\), \(\var(X)\) as:

\[
\var(X) = \frac{1}{n-1}\sum_{i=1}^{n}(X_{i}-\E[X])^{2}
\]

The formatting here is incredibly pedantic. In the right situation, it's also incredibly important.

# Try it Yourself: Formatting Environments

:::{.sol data-latex=""}
This is a solution.
:::

# Try it Yourself: Displaying Code You Don't Want to Run

```{r code=readLines("DS_S03_Example_Code.R"), eval=FALSE}
```