---
title: "Data Cleaning: Eurovision"
author: "Max Glonek"
date: "Trimester `r if(as.integer(format(Sys.time(),'%m'))<5){1}else if(as.integer(format(Sys.time(),'%m'))<9){2}else{3}`, `r format(Sys.time(),'%Y')`"
output: 
  html_document:
    number_sections: yes
    df_print: paged
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  fig.width = 6, 
  fig.asp = 0.618, 
  out.width = "70%",
  fig.align = "center", 
  root.dir = '../'
)
pacman::p_load(tidyverse, tidymodels)
theme_set(theme_bw())
if (Sys.info()[["sysname"]]=="Windows") { # if windows
  wdpath <- glue::glue("C:/Users/maxgl/Box Sync/DS_{format(Sys.time(),'%y')}_T{if(as.integer(format(Sys.time(),'%m'))<5){1}else if(as.integer(format(Sys.time(),'%m'))<9){2}else{3}}")
} else if (Sys.info()[["sysname"]]=="Darwin") { # if mac
  wdpath <- glue::glue("~/Box Sync/DS_{format(Sys.time(),'%y')}_T{if(as.integer(format(Sys.time(),'%m'))<5){1}else if(as.integer(format(Sys.time(),'%m'))<9){2}else{3}}")
}
```

# Rule 1: Look at the Data

# Rule 2: Is There a Package to Deal With That

```{r}
songs <- jsonlite::fromJSON(glue::glue("./data/eurovision-lyrics.json"))
```

# Rule 3: Break it Into Pieces

```{r}
songs[[1]]
```

```{r}
parse_song <- function(song){
  country <- song$Country
  artist <- song$Artist
  title <- song$Song
  year <- song$Year
  lyrics <- song$Lyrics
  translation <- song$`Lyrics translation`
  place <- song$Pl.
  tibble(
    country, artist, title, year, lyrics, translation, place
  )
}
parse_song(songs[[1]])
```

# Rule 4: Get it Into a Tibble

```{r}
songs <- 
  songs %>% 
  map_df(parse_song, .id = "ID")
songs
```

# Rule 5: Look at Each Column

We need to make sure all of our data makes sense in context. You can find the source data at this link: <https://www.kaggle.com/minitree/eurovision-song-lyrics>

## ID

This is an ID number assigned to each song. We should check that the song IDs are all unique.

```{r}
songs %>% count(ID) %>% filter(n > 1)
```

They are unique, but they're also not useful, so delete them.

```{r}
songs <- 
  songs %>% 
  select(-ID)
songs
```

## Artist

Are there any artists who appear more than once?

```{r}
songs %>% 
  count(artist) %>%
  arrange(-n)
```

```{r}
songs %>%
  filter(str_detect(artist, "Jed"))
```

```{r}
songs %>%
  filter(country == "Australia")
```

## Title

```{r}
songs %>%
  count(title) %>% 
  arrange(-n)
```

## Year

```{r}
range(songs$year)
```

```{r}
songs %>% 
  count(year) %>% 
  ggplot(aes(year, n)) + 
  geom_point() + 
  theme(axis.text.x = element_text(angle = -90, hjust=0))
```

## Lyrics

```{r}
songs %>% 
  sample_n(5) %>% 
  pull(lyrics) %>% 
  cat(sep = "\n\n\n")
```

## Translation

```{r}
songs %>% 
  sample_n(5) %>% 
  pull(translation) %>% 
  cat(sep = "\n\n\n")
```

## Place

```{r}
songs %>% count(place)
```

```{r}
length(unique(songs$year))
```

```{r}
songs %>% 
  count(year, place) %>% 
  filter(place == "1") %>% 
  arrange(-n)
```

According to Wikpedia, there were four winners in 1969:

<https://en.wikipedia.org/wiki/Eurovision_Song_Contest>

> On only one occasion have multiple winners been declared in a single contest: in 1969, four countries finished the contest with an equal number of votes and due to the lack of a tie-break rule at the time, all four countries were declared winners.[7][157] 

# Rule 6: You May Need to Go Back

Our original parse function just grabbed `translation` as is. After looking at `translation`, we can see that this variable says `English` if the original song lyrics are in English. English.

# Rule 7: Create New Columns?

## Decade

```{r}
songs
songs <- 
  songs %>% 
  mutate(
    year = parse_number(year), 
    decade = year - year %% 10
  )
songs %>% 
  ggplot(aes(year, decade)) + 
  geom_point()
```

## Lyrics

```{r}
songs %>% 
  count(translation) %>%
  arrange(-n)
```

```{r}
songs <- 
  songs %>% 
  mutate(
    english_lyrics = ifelse(
      translation == "English",
      lyrics, translation
    ) 
  )
```

# Rule 7: Save the Data and Write it Up

It's critically important to make detailed notes about any changes you have made when cleaning the data.

```{r}
filenamepath <- glue::glue("./data/{lubridate::today()}-eurovision.rds")
filenamepath
write_rds(songs,filenamepath)
```